---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Impacts of Land Use on Water Quality in Minnesota"
subtitle: "https://github.com/lhr12/HDA_Project"
author: "Keith Bollt, Jake Greif, Felipe Raby-Amadori, & Lindsay Roth"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

<Arrow brackets are used for annotating the RMarkdown files. Text within these brackets should not appear in the final version of the PDF document>

<**General Guidelines**>
<1. Write in scientific style> 
<2. [Global options for R chunks](https://rmarkdown.rstudio.com/lesson-3.html) should be set so that only relevant output is displayed>
<3. Make sure your final knitted PDF looks professional. Format tables appropriately, size figures appropriately, make sure bulleted and numbered lists appear as such, avoid awkwardly placed page breaks, etc.>

```{r setup, include=FALSE}
# Set your working directory
getwd()

# Load your packages
suppressMessages(library(tidyverse))
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(viridis)
library(knitr)
library(lubridate)
library(lme4)
library(nlme)
library(trend)
library(gridExtra)
library(corrplot)
library(car)
library(bestNormalize)
library(stats)
library(LAGOSNE)
library(sf)
library(maps)
library(mapview)
library(lme4)

# Set your ggplot theme
theme_set(theme_classic())

# Load your datasets
#Early
LAGOS.MN.processed.Early <- read.csv("./data/processed/LAGOS.MN.processed.Early.csv")
#Late
LAGOS.MN.processed.Late <-read.csv("./data/processed/LAGOS.MN.processed.Late.csv")
#Prime
LAGOS.MN.processed.Prime <-read.csv("./data/processed/LAGOS.MN.processed.Prime.csv")
```


# Rationale and Research Questions

<Write 1-2 paragraph(s) detailing the rationale for your study. This should include both the context of the topic as well as a rationale for your choice of dataset (reason for location, variables, etc.) A few citations should be included to give context for your topic. You may choose to configure autoreferencing for your citations or add these manually.>

<At the end of your rationale, introduce a numbered list of your questions (or an overarching question and sub-questions). Each question should be accompanied by one or more working hypotheses, inserted beneath each question.>

Land use has a large impact on nutrient runoff into streams, lakes, and other water bodies. Understanding the causes of nutrient problems will better inform management strategies. A study looking at land use and water quality in the Chaohu Basin in China found that forest and grassland has a statistically significant negative correlation with water quality, while urbanization has a statistically significant positive correlation with water quality and farmland has a more complex relationship (Huang et al., 2013). We applied some of the basic research questions of the Huang et al. study to the US, where Nutrient management has been a challenge for states in the effort to control harmful algal blooms and coastal dead zones (Wurtsbaugh et al., 2019). We selected the state of Minnesota for our study because it has wide variety of land uses including a large urban center (the Twin Cities metropolitan area), natural lands in the northern part of the state, and large agricultural areas. It also has a large number of lakes with water quality sampling. In addition, studies performed in Minnesota have shown that water quality has been adversely affected by land use changes (Christensen et al., 2012); (Maalim et al., 2013). There may also be seasonal relationships between water quality and land use; for example, we know that farmers apply nutrients at certain times of year and lakes naturally have higher levels of algae in summer months. It is important that states like Minnesota have sound nutrient management regimes because they affect water quality downstream of their borders. The Gulf of Mexico has been experiencing hypoxia, or low dissolved oxygen levels, as a result of agricultural runoff in the Mississippi River basin. Minnesota is the source of the Mississippi River and a large agricultural state, so it is important that Minnesota properly manages its agricultural runoff (Minnesota Pollution Control Agency, 2019).


* * 
Research Questions:
	In order to evaluate the relationship between land use wand lacustrine water quality in the state of Minnesota, we asked the following two questions to help guide our research:
1.	What are the predictors of nutrients based on land use in watersheds in the state of Minnesota?
Hypothesis 1a: The percentage of urbanized land in a lake’s watershed in the state of Minnesota will have a statistically significant negative effect on water quality, as measured by higher Chlorophyll a concentrations and lower secchi depth measurements.
Hypothesis 1b: The percentage of agricultural land in a lake’s watershed in the state of Minnesota will have a statistically significant negative effect on water quality, as measured by higher Chlorophyll a concentrations and lower secchi depth measurements.
2.	How do you characterize seasonal variation between the predictors of nutrients?
Hypothesis 2: The water quality in Minnesota lakes will demonstrate a statistically significant seasonal pattern.

Goals:
	By testing our hypotheses, we aim to accomplish the following goals:
1.	Determine how land use, watershed size, and ecoregion explain variation in nutrient loading indicators.
2.	Discern whether there a seasonal trends in nutrient loading indicators based on land use types, watershed size, and ecoregion.
3.	Provide insight to inform decisions about nutrient management practices based on land use types, watershed size, and ecoregion.


\newpage

# Dataset Information

<Provide information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the metadata file for the dataset but formatted in a way that is more narrative.>

<Describe how your team wrangled your dataset in a format similar to a methods section of a journal article.>

<Add a table that summarizes your data structure (variables, units, ranges and/or central tendencies, data source if multiple are used, etc.). This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.>

The data used in this analysis include data from the Lake Multi-Scaled Geospatial and Temporal Database (LAGOSNE) and the EPA ecoregion spatial datasets.


LAGOSNE is a collection of several data modules that contain information on lakes in the northern United States. The modules contain data from thousands of lakes in 17 states in the northeastern and midwestern United States, from Missouri to Maine. The dataset includes a complete list of all lakes bigger than 4 hectacres in the 17 state area, and water quality data on a large number of lakes, spanning every state. 

Ecoregions are developed by the US Environmental Protection Agency and are used by planning managers to understand the type of land use that occurs in different regions of the United States. There are different levels of ecoregions. Level I has low resolution, and divides North America into 15 ecological regions, while Level IV offers fine ecological resolution for each state. This data was published by the U.S. EPA Office of Research and Development (ORD) - National Health and Environmental Effects Research Laboratory (NHEERL). Minnesota contains parts of three distinct Level II ecoregions, which we selected for our project. This level of resolution allows us to broadly capture the natural ecosystem variation in different parts of the state in our statistical models with sufficient water quality data per ecoregion, conditions that could not be met with higher resolution ecoregions.


```{r ,echo=FALSE}
Variables.table <- read.csv("./Output/Variable_Descriptions_Table.csv")
library(knitr)
kable(Variables.table)
```


\newpage

# Exploratory Analysis 

<Insert exploratory visualizations of your dataset. This may include, but is not limited to, graphs illustrating the distributions of variables of interest and/or maps of the spatial context of your dataset. Format your R chunks so that graphs are displayed but code is not displayed. Accompany these graphs with text sections that describe the visualizations and provide context for further analyses.>

<Each figure should be accompanied by a caption, and each figure should be referenced within the text>

We came in to our exploratory analysis knowing we wanted to look at how water quality was affected by varying land use. Our exploratory analysis had four main goals: selecting a study area, selecting a study period, selecting what land use classifications (independent variables) and water quality proxies (dependent variables) to study, and preparing a dataframe to test our research questions and hypotheses.
	Our first step was to load the LAGOSNE data columns we though we would need and to look at how much data each state in our study contained. We decided on Minnesota because of its high number of lakes in the LAGOSNE dataset, and its high land use variation. We observed that the latest land use classification by USGS was from 2011; the second oldest was from 2006. Therefore, we limited our study to the period from 2009 through 2015 inclusive, so we did not extrapolate too much from our 2011 land use dataset. 
	We filtered out just the variables we wanted for our study and created a “LAGOS.MN.processed” dataframe. We decided to use Chlorophyll a and secchi depth as proxies of water quality after we visualized our data and confirmed that we had enough data on these two variables in Minnesota between 2009 and 2015 to make claims about land use and water quality; in fact we had over 19,000 lakes with data. We decided what land uses we cared about, based on our background readings and prior knowledge of water quality and land use, grouped some land uses together, and assigned them names in our dataframe. We also created an IWS ratio column, which is the lake area divided by watershed area. We reasoned a big lake in a small watershed would be affected by land use differently than a small lake in a big watershed, and we wanted this to be captured in our analysis.
	Our next step was to create a growing “seasons” column in order to account for seasonality in our statistical models. We divided up the calendar year into three bins: Early (May 15 and Before), Prime (May 16 - September 30), and Late (October 1 and later). These season designations are somewhat arbitrary, but are based off of our prior knowledge of when the prime biological growing season is in Minnesota. We created a dataframe for each of the three seasons. 
	At this point, we visualized land use percentage versus secchi depth and versus Chlorophyll a for three representative land uses for the Late season. We were hoping to visually confirm that we might find a trend when we ran our statistical tests in our analysis section. We found that some ggplots showed a possible trend, whereas some ggplots did not appear to show a trend. This told us that we had, in fact selected useful land use classifications and useful seasonal demarcations. We performed a summary function, and gave all of our columns useful names and removed NAs. Our statistical tests would not all run if our dataset contained NAs.The next step we performed was to convert our dataframes to sf dataframes so we could add Level II ecoregion column.
	Our final exploratory analysis test was to see if we had enough lake data for each Level II ecoregion for each of our three seasons. We created three tables, one for each season, with a lake sample count for each ecoregion. From these tables, we learned that we had enough data to proceed with our statistical analysis, and therefore Level II ecoregions were the correct resolution for our study. We also displayed this information spatially. We discovered that there are a large number of lake samples around the Twin Cities area, and that there is enough spatial variation in our data to capture the three different Level II ecoregions for all three of our seasons. We concluded our data wrangling by exporting our three non-sf season processed dataframes as csv files to our processed data folder.


\newpage

# Analysis

<Insert visualizations and text describing your main analyses. Format your R chunks so that graphs are displayed but code and other output is not displayed. Instead, describe the results of any statistical tests in the main text (e.g., "Variable x was significantly different among y groups (ANOVA; df = 300, F = 5.55, p < 0.0001)"). Each paragraph, accompanied by one or more visualizations, should describe the major findings and how they relate to the question and hypotheses. Divide this section into subsections, one for each research question.>

<Each figure should be accompanied by a caption, and each figure should be referenced within the text>

## Question 1: <insert specific question here and add additional subsections for additional questions below>

## Question 2: 




\newpage

# Summary and Conclusions
<Summarize your major findings from your analyses in a few paragraphs. What conclusions do you draw from your findings? Relate your findings back to the original research questions and rationale.>

\newpage

# References
<add references here> 
Christensen, Victoria G., et al. "Relations between Retired Agricultural Land, Water Quality, and	Aquatic-Community Health, Minnesota River Basin." Journal of Environmental Quality,	vol. 41, no. 5, 2012, pp. 1459-72. ProQuest,	https://login.proxy.lib.duke.edu/login?url=https://search.proquest.com/docview/1	95114913?accountid=10598.

Huang, J., Zhan, J., Yan, H., Wu, F., Xiangzheng Deng, X. “Evaluation of the Impacts of Land	Use on Water Quality: A Case Study in The Chaohu Lake Basin,” The Scientific World	Journal, vol. 2013, Article ID 329187, 7 pages,	2013. https://doi.org/10.1155/2013/329187.

Maalim, Fukhrudin K., Melesse, Assefa M., Belmont, Patrick., Gran, Karen B. "Modeling the	impact of land use changes on runoff and sediment yield in the Le Sueur watershed,	Minnesota using GeoWEPP." CATENA Volume 107, 2013, Pages 35	45, https://doi.org/10.1016/j.catena.2013.03.004.

“Nutrient Reduction Strategy.” Minnesota Pollution Control Agency, 6 Mar. 2019, www.pca.state.mn.us/water/nutrient-reduction-strategy#resources-b7ca91dc.

Ramstack, Joy M., Sherilyn C. Fritz, and Daniel R. Engstrom. "Twentieth Century Water Quality	Trends	in Minnesota Lakes Compared with Presettlement Variability." Canadian Journal	of Fisheries and Aquatic Sciences, vol. 61, no. 4, 2004, pp. 561-576.

